---
title: "Medical Cost LM"
output: html_notebook
---

```{r}
#Read CSV into R
MedicalCostData <- read.csv("MedicalCost.csv", header=TRUE)

#Display the first few elements of the data to make sure the data is read
head(MedicalCostData)
```

```{r}
#Subset of the data
charges <- MedicalCostData$charges
age <- MedicalCostData$age

#Plotting age against charges to see if there is a correlation
plot(age, charges, pch=20, col="red")
```

```{r}
#BMI subset eof the data
bmi <- MedicalCostData$bmi

#plotting bmi
plot(bmi, charges, pch=20, col="red")
```

```{r}
#Convert the sex variable in a binary variable
MedicalCostData$sex <- factor(MedicalCostData$sex, levels = c("male", "female"), labels = c(0,1))
```

```{r}
#Convert the smoker variable in a binary variable
smoker_levels <- c("no", "yes")
MedicalCostData$smoker <- factor(MedicalCostData$smoker, levels = smoker_levels, label = c(0,1))
```

```{r}
#Delete the region data because it is not significant to our regression
MedicalCostData$region <- NULL
MedicalCostData
```
```{r}
#Turning the sex and smoker variables into numerical variables
#cor(MedicalCostData[ , ])
MedicalCostData$sex = as.numeric(MedicalCostData$sex) - 1
MedicalCostData$smoker = as.numeric(MedicalCostData$smoker) - 1
cor(MedicalCostData[ , ])
```

```{r}
#Create a regression model
lmCharges = lm(charges~age + sex + bmi + children + smoker, data = MedicalCostData)
summary(lmCharges)
```
The sex variable is statistically insignificant.

```{r}
#Re-run the regression model without the sex variable
lmCharges = lm(charges~age + bmi + children + smoker, data = MedicalCostData)
summary(lmCharges)
```
The predictor with the largest impact on charges would be smoker because it has the largest coefficient.

The r-squared value is high, showing that the data highly fits the model. The adjusted R-squared changed very slightly from before the insignificant predictor was dropped, showing that there insignificant predictor had almost no impact on the model.

```{r}
#Split the data into a training set and testing set

#Set the split to 80%
num_samples = dim(pumpkinsData) [1]
sampling.rate=0.8

training <- sample(1:num_samples, sampling.rate * num_samples, replace=FALSE)
trainingSet <- subset(MedicalCostData[training, ])

testing <- setdiff(1:num_samples, training)
testingSet <- subset(MedicalCostData[testing, ])

```

```{r}
#Fit a model for the training set
lmCharges = lm(charges~age + bmi + children + smoker, data = TrainingSet)
summary(lmCharges)
```

```{r}
#Perform predictions 
model <- lm(charges ~ ., TestingSet)
p <- predict(model, TestingSet)
```

```{r}
#Evaluate the model by calculating the error
error = p - TestingSet
mse = mean(error^2)
```
